---
title: "Statistical_Infer_Part1.Rmd"
author: "Joe Okelly"
output: html_document
---


## Overview 

In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.


## Simulations 

The exponential distribution can be simulated in R with `rexp(n, lambda)`, where `lambda` is the rate parameter and `n` is the number of Simulations. 'lambda' is set to `0.2`.
load the `ggplot2` plotting library.

```{r}
library(ggplot2)
library(knitr)
```

Intialiaze the variables 

```{r warning=FALSE, error=FALSE}
num_Simulation <- 1000
sampSize <- 40
lambda <- 0.2
set.seed(300413)

```

Define a matrix of 1000 rows x 40 columns, corresponds to Number of Sumulations and Sample Size.

```{r warning=FALSE, error=FALSE}

simu_data <- replicate(num_Simulation, rexp(sampSize, lambda))
mean_simulation_data <- apply(simu_data, 2, mean)

simu_Matrix <- matrix(rexp(n = num_Simulation * sampSize, rate = lambda), num_Simulation, sampSize)
simu_Mean <- rowMeans(simu_Matrix)
simu_Data <- data.frame(cbind(simu_Matrix, simu_Mean))



```

Define a ggplot to visiualise the data.

```{r}

ggplot(data = simu_Data, aes(x = simu_Mean)) + 
  geom_histogram(aes(y = after_stat(density)), binwidth = 0.3, fill = "lightgreen", color = "black") + 
  labs(title = "Mean Distribution", x = "Simulated Means ", y = "Density") +
  geom_vline(aes(xintercept=mean(simu_Mean)), color="red", linetype="dashed", linewidth=1)
  
```

```{r}
actual_Mean <- mean(simu_Mean)
theo_Mean <- (1 / lambda)
act_Variance <- var(simu_Mean)
theo_Variance <- ((1 / lambda) ^ 2) / sampSize

print(paste("Actutal Mean :", actual_Mean))
print(paste("Theo Mean :", theo_Mean))
print(paste("Actual VAriance :", act_Variance))
print(paste("Theorical VAriance :", theo_Variance))

```

```{r}

qqnorm(mean_simulation_data)
qqline(mean_simulation_data, col = "magenta")

```

